### 2025-12-01 Paper/main.tex
camera-ready

\documentclass[sigplan,screen]{acmart}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
\usepackage{stmaryrd}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{intersections}
\usetikzlibrary{automata}
\usetikzlibrary{positioning}
\usepackage{booktabs}
%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
\usepackage{listings}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{balance}
\usepackage{microtype}

\input{comm}

\lstdefinelanguage{SMTLIB}{
  morekeywords={set-logic, declare-fun, assert, check-sat, get-model},
  sensitive=true,
  morecomment=[l];,
  morestring=[b]"
}

\lstdefinelanguage{Isabelle}{
    keywords={definition,theorem,lemma,proof,qed,by,auto,simp,apply,done,
              where,fixes,assumes,shows,if,then,else,let,in,case,of,
              record,datatype,type_synonym,fun,function,primrec},
    sensitive=true,
    morecomment=[s]{(*}{*)},
    morestring=[b]",
  }

\lstset{
  basicstyle=\ttfamily,
  keywordstyle=\color{blue}\bfseries,
  morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and,locale,set-logic,declare-fun,assert,check-sat,get-model},
  escapeinside={|}{|},
  sensitive=true,
  commentstyle=\color{green!50!black},
  morecomment=[l];,
  morestring=[b]",
  breaklines=true,
  showstringspaces=false
}

\bibliographystyle{ACM-Reference-Format}% the mandatory bibstyle

%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title[Certified Symbolic Finite Transducers]{Certified Symbolic Finite Transducers: Formalization and Applications to
String Analysis}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\author{Shuanglong Kan}
\orcid{0000-0002-4132-2626}
\affiliation{%
  \department{ Verified System Design Automation Group}
  \institution{Barkhausen Institut}
  \city{Dresden}
  \country{Germany}
}
\email{shuanglong.kan@barkhauseninstitut.org}

\author{Anthony W. Lin}
\orcid{0000-0003-4715-5096}
\affiliation{
  \department{Department of Computer Science}
  \institution{University of Kaiserslautern-Landau \& MPI-SWS}
  \city{Kaiserslautern}
  \country{Germany}
}
\email{awlin@mpi-sws.org}

\input{abstract.tex}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10003752.10003790.10003794</concept_id>
       <concept_desc>Theory of computation~Automated reasoning</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003766.10003773.10003774</concept_id>
       <concept_desc>Theory of computation~Transducers</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Automated reasoning}
\ccsdesc[500]{Theory of computation~Transducers}

\keywords{symbolic transducers, Isabelle/HOL, string theory, SMT solvers} %TODO mandatory; please add comma-separated list of keywords
\maketitle

\input{introduction.tex}

\input{motivation.tex}

\input{SFT.tex}

\input{formalization.tex}

\input{application.tex}

\input{relatedwork.tex}

\input{conclusion.tex}

\balance
\bibliography{literature}
\end{document}